---
title: "ProSportsDraftDataVignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ProSportsDraftDataVignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::install_github("Ginsburg1/ProSportsDraftData")
library(ProSportsDraftData)
```


## Available Datasets

The `ProSportsDraftData` package provides several datasets:

- **nfl_data**: NFL draft data
- **nhl_data**: NHL draft data
- **nba_data**: NBA draft data

## The `nfl` Dataset

The `nfl` dataset includes the following datapoints:

- **source**: The source of the data.
- **name**: The name of the player.
- **year**: The year of the draft.
- **rank**: The rank of the player.
- **round**: The round in which the player was drafted.
- **height**: The height of the player.
- **weight**: The weight of the player.
- **position**: The position of the player.
- **college**: The college the player attended.
- **pros**: The pros of the player’s abilities.
- **cons**: The cons of the player’s abilities.
- **similar_player**: A similar player for comparison.
- **similar_player_bio**: Similar player comparison.
- **summary**: A summary of the player’s abilities.
- **arm_length**: The arm length of the player.
- **hand_length**: The hand length of the player.
- **next_gen_production_score**: The Next Gen production score.
- **next_gen_athleticism_score**: The Next Gen athleticism score.
- **forty_yard_dash**: The forty-yard dash time.
- **vertical_jump**: The vertical jump height.
- **nfl_prospect_grade**: The NFL prospect grade.
- **home_town**: The hometown of the player.
- **broad_jump**: The broad jump distance.
- **three_cone_drill**: The three-cone drill time.
- **twenty_yard_shuttle**: The twenty-yard shuttle time.
- **bench_press**: The bench press reps.
- **college_abbrivation**: The abbreviation of the college.
- **pre_draft**: Pre-draft information.
- **post_draft**: Post-draft information.
- **position_rank**: The position rank of the player.
- **overall_rank**: The overall rank of the player.
- **grade**: The grade of the player.
- **school**: The school the player attended.
- **yds**: Yards the player ran.
- **ypa**: The yards per attempt.
- **ypr**: The yards per reception.
- **tds**: Number of touchdowns the player performed.
- **ints**: The interceptions.
- **rtg**: The rating of the player.
- **tkls**: The number of tackles of the player.
- **tfl**: The tackles for loss.
- **ypc**: The yards per carry.
- **pbu**: The pass break-ups of the player.
- **twenty_plus**: Plays of twenty or more yards.
- **sacks**: Number of sacks of the player.
- **gms**: The number of games played.
- **strts**: The number of games started.
- **sk_all**: The number of sacks allowed.
- **age**: The age of the player.
- **main_selling_point**: The main selling point of the player.
- **description**: The description of the player.
- **scouting_report**: The scouting report.
- **score**: Player’s score from 1-100.

#Loading NFL Data
First, let's load the NHL data provided by the package.

```{r}
head(nfl_data)
```

## The `nhl_df` Dataset

The `nhl_df` dataset includes the following datapoints:

- **source**: The source of the data.
- **name**: The name of the player.
- **year**: The year of the draft.
- **rank**: The rank of the player.
- **round**: The round in which the player was drafted.
- **date_of_birth**: The date of birth of the player.
- **age**: The age of the player.
- **current_team**: The current team of the player.
- **league**: The league in which the player plays.
- **scouting_report**: The scouting report of the player.
- **team_fit**: The fit of the player within the team.
- **likely_to_play**: The likelihood of the player to play in the NHL.
- **position**: The position of the player.
- **city**: The city associated with the player.
- **team**: The team of the player.
- **analysis**: An analysis of the player’s abilities.


#Loading NHL Data
First, let's load the NHL data provided by the package.

```{r}
head(nhl_data)
```

#Filtering By Data Source
The NHL data has been compiled from multiple sources. The three different datasouces are

- **Base (ESPN.com)**: Data from ESPN.com
- **ESPN**: Data from ESPN
- **NHL.com**: Data from NHL.com

To see the data from a specific source, use of the following functions:

```{r}
nhl_base_data <- nhl_data_base()
head(nhl_base_data)

nhl_data_espn()
nhl_data_nhl.com()
```


## The `nba_csv` Dataset

The `nba_csv` dataset includes the following datapoints:

- **source**: The source of the data.
- **name**: The name of the player.
- **year**: The year of the draft.
- **rank**: The rank of the player.
- **round**: The round in which the player was drafted.
- **height**: The height of the player.
- **weight**: The weight of the player.
- **school**: The school the player attended.
- **college_abbreviation**: The abbreviation of the college.
- **pros**: The pros of the player’s abilities.
- **cons**: The cons of the player’s abilities.
- **verdict**: The verdict on the player's potential.
- **pre_draft_analysis**: Pre-draft analysis of the player.
- **post_draft_analysis**: Post-draft analysis of the player.
- **ranking**: The player's ranking.
- **player_score**: The player's score.
- **position**: The position of the player.
- **grade**: The grade of the player.
- **pts_total**: Total points scored by the player.
- **pts_per40**: Points scored per 40 minutes.
- **reb_total**: Total rebounds by the player.
- **reb_per40**: Rebounds per 40 minutes.
- **ast_total**: Total assists by the player.
- **ast_per40**: Assists per 40 minutes.
- **efg**: Effective field goal percentage.
- **fga**: Field goals attempted.
- **stl_total**: Total steals by the player.
- **stl_per40**: Steals per 40 minutes.
- **blk_total**: Total blocks by the player.
- **blk_per40**: Blocks per 40 minutes.
- **tpt**: Three-point shots made.
- **tpa**: Three-point shots attempted.
- **ft**: Free throws made.
- **fta**: Free throws attempted.
- **ts**: True shooting percentage.
- **ts_att**: True shooting attempts.
- **main_selling_point**: The main selling point of the player.
- **age**: The age of the player.
- **wingspan**: The wingspan of the player.
- **standing_reach**: The standing reach of the player.
- **similar_player**: A similar player for comparison.
- **profile**: The profile of the player.
- **analysis**: An analysis of the player’s abilities.
- **projection**: The projection of the player's future performance.
- **date_of_birth**: The date of birth of the player.
- **hometown**: The hometown of the player.
- **high_school**: The high school the player attended.
- **rebounding**: The rebounding ability of the player.
- **potential**: The potential of the player.
- **post_skills**: The post skills of the player.
- **intangibles**: The intangibles of the player.
- **international_team**: The international team of the player.
- **athleticism**: The athleticism of the player.
- **size**: The size of the player.
- **defense**: The defensive skills of the player.
- **strength**: The strength of the player.
- **quickness**: The quickness of the player.
- **leadership**: The leadership qualities of the player.
- **jumpshot**: The jump shot ability of the player.
- **nba_ready**: The NBA readiness of the player.
- **mock_draft_rank**: The player's mock draft rank.
- **big_board_rank**: The player's big board rank.
- **overall_rank**: The overall rank of the player.


#Loading NBA Data
First, let's load the NBA data provided by the package.

```{r}
head(nba_data)
```
