---
editor_options: 
  markdown: 
    wrap: 72
---
## `ProSportsDraftData: An R package for Professional Sports Draft Data`

# Summary

The`ProSportsDraftData` package provides researchers and analysts with a tool to view and analyze scouting reports for potential draft picks in three major men's professional sport drafts (NBA, NFL, and NHL). Professional teams are alloted draft slots where they select amateur players to join their team.

This package provides scouting reports from 13 sources for the three leagues. The data allows for many potential analyses, including: determining if there are biases in scouting reports for athletes of different racial or ethnic backgrounds and improving draft strategies for professional teams. The significance of sports drafts is highlighted by events such as quarterback Brock Purdy being selected with the last pick in the 2022 NFL draft but ending up as one of the most successful quarterbacks in the league. The scouting report data provided in this package could offer reasons why he was drafted so late.


# Statement of Need

The draft is an important element for all professional sports leagues for both fans and the teamss themselve. The `ProSportsDraftData` package is an R package that combines all the different draft data into one source allowing easier research. This package enables a deeper investigation of the draft, offering insights that may have not been analyzed before. 

This package is further important due to the popularity of professional sports. For instance, the 2024 NFL draft took place in Detroit with over 700,000 live spectators and the 2024 NHL draft was at the popular Las Vegas Sphere. This package fills a current lack of simple scouting report data accessibility. It allows users to more easily view and analyze data about draft picks. R is an important platform for this package because of its open source nature and use in the sports analytics community, see Elmore and Urb book citation.

In a 2016 article, Michael Shuckers CITE discusses the use of historical data, such as player demographics, pre-draft performance, and scouting rankings, to build a statistical prediction model for NHL players' performance. The model was tested on two sets of players, consistently showing the statistical approach to be more effective at predicting player success than traditional draft methods. The `ProSportsDraftData` R package would greatly aid such an analysis because it would the data gathering process easier and give researchers a wider dataset to draw from.


## Data

- How the scraping was carried out

To scrape the data, a combination of RSelenium and RVest R packages were used. Whenever a website required manipulating (e.g. clicking, changing website dimensions) RSelenium was used, otherwise RVest was used. The scraped data was organized into tibbles with corresponding player names.

- What the data sources are

The `ProSportsDraftData` package provides several data sets:

-   **nfl_data**: NFL draft data
-   **nhl_data**: NHL draft data
-   **nba_data**: NBA draft data

## Sources and Citations

-   **Base**: Base was used to establish a standard naming convention
    for players, avoiding confusion caused by nicknames or different
    variations of names among different sources.
    -   **NFL**: <https://www.espn.com/nfl/draft/rounds/_/round/>
    -   **NBA**: <https://www.espn.com/nba/draft/rounds/_/round/>
    -   **NHL**: <http://www.mynhldraft.com/>
-   **The Ringer**:
    -   **NBA**: <https://nbadraft.theringer.com/>
    -   **NFL**: <https://nfldraft.theringer.com/>
-   **ESPN**:
    -   **NFL**: <https://www.espn.com/nfl/draft/rounds/_/round/>
    -   **NBA**: <https://www.espn.com/nba/draft/rounds/_/round/>
    -   **NHL**: <https://www.espn.com/nhl>
-   **NBA.com**:
    -   **NBA**: <https://www.nba.com/draft/>
-   **NBADraft.net**:
    -   **NBA**: <https://www.nbadraft.net/players/>
-   **NBA.com**:
    -   **NHL**: <https://www.nhl.com/news/>
-   **walterfootball.com**:
    -   **NFL**: <https://walterfootball.com/>
-   **NFL.com**:
    -   **NFL**: <https://www.nfl.com/prospects/participants/all/>

# Functionality

## A few examples of using the package

Different heights of players from the ESPN source in the NFL.
```{r}
library(ProSportsDraftData)
nfl_data_espn() |>
     dplyr::filter(!is.na(height)) |>
     dplyr::group_by(height) |>
     dplyr::summarize(count = dplyr::n()) |>
     dplyr::arrange(desc(count))
```


All of the players from Alabama according to ESPN. All of the different leagues that NHL players came from according to ESPN.

```{r}
library(ProSportsDraftData)
nfl_data_espn() |>
  dplyr::filter(college == "Alabama")
```


List of "similar players" to the potential NBA draft pick from nbadraft.net that were under 74 inches.

```{r}
library(ProSportsDraftData)
nba_data_nbadraft_net() |>
  dplyr::filter(!is.na(height), height < 74) |>
  dplyr::select(name, similar_player)
```

# Conclusion

The `ProSportsDraftData` R package consolidates and analyzes draft data from the NFL, NBA, and NHL, addressing a significant gap in sports analytics by providing a convenient source for the data. This package compiles datasets from various sources, allowing researchers and analysts to view and analyze various elements of draft data. As sports drafts continue to become more important, the package offers organized information in one source.

# Citations
Schuckers, Michael E. “Raft by Numbers: Using Data and Analytics to Improve National Hockey League (NHL) Player Selection.” Sports Anlaysics Confriance, sports analytics conference, 11 Mar. 2016, myslu.stlawu.edu/\~msch/SchuckersDraftByNumbers2016.pdf.
