---
editor_options: 
  markdown: 
    wrap: 72
---
## `ProSportsDraftData: An R package for Professional Sports Draft Data`\

# Summary

The`ProSportsDraftData` package provides researchers and analysts with a tool to view and analyze scouting reports for draft picks. Through using web scraping tools in R, this package makes it easier to view biases inscouting reports and improving draft strategies. The significance of sports drafts is highlighted by events like Brock Purdy being selected 256th overall. Even though he was selected last overall in the draft,certain verbiage could be analyzed to see why teams did not draft him earlier and why he became a prominent quarterback. Another reason why this package was created is due to the popularity of the major sports. For instance, the 2024 NFL draft took place in Detroit with over 700,000 live spectators and the NHL draft was at the Las Vegas Sphere. As drafts have become more and more prominent, this package will fill a currenthole within the R that allows users to more easily view data about draft picks. This package aims to fill a gap in R, making it easier to access and analyze draft data from various sports in one source.

# Statement of Need

All the major sports in the US, the draft is an important element for all leagues for both fans and the teams. The `ProSportsDraftData` package is only R package that combines all the different draft data into one source allowing easier research. This package enables a deeper investigation into these biases and other aspects of scouting language, offering insights that have not been analyzed using an R package before. In addition, we suppose there may be biased language in these reports because in some sports, as most of the players are from an ethnic minority. One similar paper, by Michael Shuckers discusses the use of historical data, such as player demographics, pre-draft performance, and scouting rankings, to build a statistical prediction model for NHL players' performance. The model was tested on two sets of players, consistently showing that the statistical approach was more effective at predicting player success than traditional draft methods. This R package would help analyze the data because it would make it easier find the data.



Data

-How the scraping was carried out

To scrape the data, I used a combination of RSelenium and RVest. Whenever a website required manipulating (e.g. clicking, changing website dimensions) RSelenium was used, else RVest was used. The scraped data was then organized into tibbles with corresponding player names.

-What the data sources are

The `ProSportsDraftData` package provides several data sets:

-   **nfl_data**: NFL draft data
-   **nhl_data**: NHL draft data
-   **nba_data**: NBA draft data

## Sources and Citations

-   **Base**: Base was used to establish a standard naming convention
    for players, avoiding confusion caused by nicknames or different
    variations of names among different sources.
    -   **NFL**: <https://www.espn.com/nfl/draft/rounds/_/round/>
    -   **NBA**: <https://www.espn.com/nba/draft/rounds/_/round/>
    -   **NHL**: <http://www.mynhldraft.com/>
-   **The Ringer**:
    -   **NBA**: <https://nbadraft.theringer.com/>
    -   **NFL**: <https://nfldraft.theringer.com/>
-   **ESPN**:
    -   **NFL**: <https://www.espn.com/nfl/draft/rounds/_/round/>
    -   **NBA**: <https://www.espn.com/nba/draft/rounds/_/round/>
    -   **NHL**: <https://www.espn.com/nhl>
-   **NBA.com**:
    -   **NBA**: <https://www.nba.com/draft/>
-   **NBADraft.net**:
    -   **NBA**: <https://www.nbadraft.net/players/>
-   **NBA.com**:
    -   **NHL**: <https://www.nhl.com/news/>
-   **walterfootball.com**:
    -   **NFL**: <https://walterfootball.com/>
-   **NFL.com**:
    -   **NFL**: <https://www.nfl.com/prospects/participants/all/>

# Functionality

## Few examples of using the package

Different heights of players from the ESPN source in the NFL.
```{r}
library(ProSportsDraftData)
nfl_data_espn() |>
     dplyr::filter(!is.na(height)) |>
     dplyr::group_by(height) |>
     dplyr::summarize(count = dplyr::n()) |>
     dplyr::arrange(desc(count))
```


All of the players from Alabama from the ESPN source. All of the different leagues that NHL players came from according to ESPN.
```{r}
library(ProSportsDraftData)
nfl_data_espn() |>
  dplyr::filter(college == "Alabama")
```



List of Similar players in the nba form nbadraft.net that were under 74 inches.
```{r}
library(ProSportsDraftData)
nba_data_nbadraft_net() |>
  dplyr::filter(!is.na(height), height < 74) |>
  dplyr::select(name, similar_player)
```

# Conclusion

The `ProSportsDraftData` R package consolidates and analyzes draft data from the NFL, NBA, and NHL, addressing a significant gap in sportsanalytics by providing a coinvent source for all data. This package webscraping tools to compile datasets from various sources, allowingresearchers and analysts to view and analyze various elements of draftdata. As sports drafts continue to become more important, the packageoffers organized information in one source.

# Citations
Schuckers, Michael E. “Raft by Numbers: Using Data and Analytics to Improve National Hockey League (NHL) Player Selection.” Sports Anlaysics Confriance, sports analytics conference, 11 Mar. 2016, myslu.stlawu.edu/\~msch/SchuckersDraftByNumbers2016.pdf.


# Acknowledgements
